---
title: "Data Tidying"
author: "Sonia Ibarra"
date: "5/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
```

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```

##Making a pipe to filter data easier
```{r}
catch_df %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% 
  head()
```

##Keeping only columns that we care about!! This is a powerful approach! 
```{r}
catch_df <-catch_df%>%
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
head(catch_df)
```

##Converting the columns to being stacked instead of side by side (wide)
```{r}
catch_df <- catch_df %>% 
  gather(species, catch, -Region, -Year)

head(catch_df)
```

##Renaming catch column to catch in thousands
```{r}
catch_df <- catch_df %>% 
  rename(catch_thousands = catch)

head(catch_df)
```

##Changing a weird value to the correct one, so the ifelse part is set to default to true and changes a value of I = 1, and the last part of the argument "catch_thousands = as.integer(catch_thousands)" leaves those value the same if it is false
```{r}
catch_df <- catch_df %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands),
         catch_thousands = as.integer(catch_thousands))
head(catch_df, 30)
```

```{r}
catch_df <- catch_df %>%
  mutate(catch = catch_thousands * 1000)
summary(catch_df)
```

```{r}
catch_df %>%
  group_by(Species, Year) %>%
  sumarize(mean_catch = mean(catch))
```


##Joins
```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)
head(region_defs)
```

``` {r}
region_defs<-region_defs%>%
  select(code, mgmtArea)
head(region_defs)
```

##Specifying which column you are joining by when the column header hasn't a differnt name. 
```{r}
catch_joined <- left_join(catch_df, region_defs, by = c("Region" = "code") )

head(catch_joined)
```